Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-18 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
chnages in Secs. III & IV

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events are detected with sufficiently high
signal-to-noise ratio \ab{in the merger}, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown,
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we \sout{take a first step to} address this question by using an
inspiral-merger-ringdown waveform model in the effective-one-body formalism,
augmented with a parametrization of the ringdown based on an expansion on the
final black hole's spin.
%
We give a prescription on how to include \ab{in} this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. \ab{In particular, we focus on gravity} theories that deviate
from general relativity through higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet, dynamical Chern-Simons, cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against \ab{the two loudest gravitational-wave ringdown signals,
  notably, GW150914 and GW200129, and we find} that while the Einstein-dilaton-Gauss-Bonnet theory cannot be constrained with
these events, we can place upper bounds on the coupling constants of cubic-
($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order ($\ell_{\rm qEFT}
\leqslant 51.3$~km) effective-field-theories of general relativity, and
of dynamical Chern-Simons gravity ($\ell_{\rm CS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently
unconstrained by inspiral-only analyses which, however, can be constrained by
merger-ringdown studies \ab{with current gravitational-wave data}.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\ab{Having reviewed the modified gravity theories we will consider, we now present
the sequential building blocks for the waveform model we use to test these theories
against GW observations.
%
We start by reviewing the parametrized ringdown spin-expansion coefficients (ParSpec) framework in Sec.~\ref{sec:review_parspec}.
%
Next, in Sec.~\ref{sec:review_pSEOB}, we review our baseline
parametrized inspiral(--plunge)-merger-ringdown (IMR) BBH waveform
model~\cite{Brito:2018rfr,Ghosh:2021mrv} and explain how we extend
it to include the ParSpec.
%
Finally, in Sec.~\ref{sec:theory_specific_qnm}, we show how we can map
theory-specific QNM calculations in modified gravity theories onto the
free coefficients in the ParSpec framework.}
%
Ultimately, this provides us with an IMR BBH waveform model, with the ringdown
portion of the model informed by QNM calculations in specific beyond-GR
theories.

@@

\ab{A general procedure to describe deviations to the QNM frequencies $\omega_{\ell m n}^{\rm Kerr}$ and
damping times $\tau_{\ell m n}^{\rm Kerr}$ of the GR Kerr BH is to write,}
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{\ell m n} &= \omega^{\rm Kerr}_{\ell m n} \, (1 + \delta \omega_{\ell m n}), \\
\tau_{\ell m n }   &= \tau^{\rm Kerr}_{\ell m  n}   \, (1 + \delta \tau_{\ell m n }),
\end{align}
\end{subequations}
%
\ab{where $\delta\omega_{\ell m n}$ and $\delta\tau_{\ell m  n}$ are the fractional deformation parameters,
$l$ and $m$ are the multipole indices, and $n$ the overtone number.}
%
This type of parametrization was adopted, for instance, in Refs.~\cite{Gossan:2011ha,Meidam:2014jpa,Brito:2018rfr,Carullo:2018sfu,Ghosh:2021mrv}.

The current LIGO-Virgo-Kagra tests of BH ringdown (see Sec.VII.A
of Ref.~\cite{LIGOScientific:2020tif} or Sec.VIII.A
of Ref.~\cite{LIGOScientific:2021sio}) take a restrictive theory-agnostic
approach towards the inference of $\delta\omega_{\ell m n}$ and
$\delta\tau_{\ell m n}$. These deviations are either assumed to occur
identically across all observed sources or belong to a generic
underlying Gaussian population. A more realistic scenario, however, is
that these parameters depend on the source BH's mass and spin.
%
Ideally, one would like to explicitly reinstate this dependence, by
%
(i) introducing deformation parameters which can be determined, once and for
all, from a specific gravity theory (GR included), and
%
(ii) making it simpler to combine constraints coming from multiple (independently observed) sources.

\ab{The ParSpec framework was introduced in Ref.~\cite{Maselli:2019mjd}. It is} an observable-based
bivariate expansion of Eq.~\eqref{eq:general_deviation}, given by
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{\ell m n} &= \frac{1}{M_{\rm f}} \sum_{j = 0}^{N_{\rm max}} \, \chi_{\rm f}^{j} \omega^{(j)}_{\ell m n} \, \left( 1 + \gamma\, \delta \omega^{(j)}_{\ell m n} \right), \\
\tau_{\ell m n}   &= M_{\rm f}     \sum_{j = 0}^{N_{\rm max}} \, \chi_{\rm f}^{j} \tau^{(j)}_{\ell m n}   \, \left( 1 + \gamma\, \delta \tau^{(j)}_{\ell m n} \right),
\end{align}
\end{subequations}
%
\ab{where $M_{\rm f}$ and $\chi_{\rm f}$ are the detector-frame final mass and spin, respectively;}
$\omega_{\ell m n}^{(j)}$ and $\tau_{\ell m n}^{(j)}$ are dimensionless coefficients of the
expansion in spin for Kerr QNMs; $\delta\omega_{\ell m n}^{(j)}$ and $\delta\tau_{\ell m n}^{(j)}$ are source-independent
dimensionless coefficients that characterize the corrections to the Kerr QNM at
each spin-order,
%
and $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence for a given modified gravity theory is contained in the
dimensionless parameter $\gamma$, which reads
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm f}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M_{\rm f}}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on \ab{the length-scale parameter $\ell$ (non-GR modifications will become
important at distances $\lesssim \ell$) and
the exponent $p$ which is related to how the beyond-GR modifications are
added to the Einstein-Hilbert action.}
%
In Eq.~\eqref{eq:def_gamma} we made \sout{$\ell$} \ag{$\gamma$} dimensionless  by using the other
length-scale associated with the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$\footnote{\ag{Throughout this paper, we maintain the convention of
using the subscript ``s" to denote source-frame quantities and plain baseline
symbols for detector-frame measurements.}} which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}. \ag{Also, dividing by the factor $G/c^2$ allows us to
express $\ell$ in physically intuitive metric units.}

In principle, a modification to GR would also affect $M$ and $\chi$ and the
expansion should be written in terms of the non-GR mass and spin, say $\bar{M}$ and
$\bar{\chi}$. If we assume that the beyond-GR corrections are included perturbatively
(as is the case of all theories described in Sec.~\ref{sec:review_theories}),
the modifications to the BH mass and spin can be absorbed into the deviations
parameters $\delta\omega$ and $\delta\tau$.
%
This means we can identify the $M$ and $\chi$ with their \emph{corresponding GR values}
\footnote{\ag{For a more detailed discussion, see Appendix A in ~\cite{Maselli:2019mjd}.}}.
%
We will see that in Sec.~\ref{sec:pe} that this assumption is indeed
satisfied in our parameter estimation studies (cf.~Fig.~\ref{fig:corner_plot}).

Finally, we remark that in the GR limit ($\gamma \to 0$)
the series~\eqref{eq:kerr_expansion} truncated at $N_{\rm max} = 4$, reproduces with 1\% accuracy
the Kerr QNMs for spins $\chi \lesssim 0.7$.
%
The values of the fitting coefficients $\omega_{k}^{(n)}$ and $\tau_{k}^{(n)}$
can be found in Ref.~\cite{Maselli:2019mjd}, Table I\agcomm{This sentence would need
to be adapted if we put the GR coefficients in our current Table II.}.
%
In Ref.~\cite{Carullo:2021dui}, the fitting coefficients were calculated up to $N_{\rm max} = 9$,
which extend the validity of the spin-expansion up to $\chi \lesssim 0.99$.
%
As we will discuss in Sec.~\ref{sec:pe}, the expansion to $N_{\rm max} = 4$ is
sufficient for our purposes.

@@

We now describe the waveform model used in our paper to infer properties of a
BBH ringdown.
%
As in Refs.~\cite{Brito:2018rfr,Ghosh:2021mrv}, we use an IMR BBH waveform
model where the parameters describing the remnant object are left additionally
free and estimated directly from the data.

The GW signal from a BBH merger can be determined in GR by a unique set of
parameters $\bm{\theta}$, that includes the masses and spins of the two BHs,
$(m_1,\, m_2,\, \vec{s}_1,\, \vec{s}_2)$, the sky location determined by the
luminosity distance $d_L$, right ascension $\alpha$ and declination $\delta$,
and the orientation of the binary given by the inclination $\iota$ and polarization $\psi$
angles.
%
The set is completed by the choice of a reference time $t_0$ and phase
$\phi_0$. If we further assume that the spins of the individual BHs are
restricted to be parallel to the orbital angular momentum, we reduce the six
components of spin to just two, $(\chi_1, \chi_2)$ and our entire parameter set
from 15 to 11.
%
Let us also define some additional parameters and set some conventions that
will be useful in our analysis later, namely, the total mass
%
$M=m_1+m_2$,
%
the chirp mass
%
$\mathcal {M}=(m_{1}m_{2})^{3/5}/(m_{1}+m_{2})^{1/5}$,
%
the asymmetric mass ratio $q=m_1/m_2$, with the convention $m_1 \geqslant m_2$ (and thus $q \geqslant 1$),
and the symmetric mass ratio of the binary, $\nu = m_1m_2/(m_1+m_2)^2$.

The polarization of the GW signal (in the observer's frame) is,
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical
harmonics.
%
% --------------------------------
% HS: maybe not necessary anymore:
% \footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.}
% --------------------------------
%
As our baseline model, that is, the GR model upon which non-GR modifications
will be added, we use the computationally efficient (time-domain) multipolar
waveform model for quasicircular spin-aligned BBH mergers described in
Ref.~\cite{Mihaylov:2021bpf} (henceforth referred to as \SEOB{}\footnote{This waveform
model is available in \texttt{LALSuite}~\cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA}
waveform approximant.}) which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$,
and $(5,5)$~\cite{Cotesta:2018fcv,Mihaylov:2021bpf}.
%
The model uses an effective-one-body (EOB) approach combined with a
post-adiabatic solution of the equations of motion to describe the
inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$.
%
An accurate description of the merger is incorporated through calibration with
NR simulations, as described in Ref.~\cite{Cotesta:2018fcv}, along with
information for the merger and ringdown phases, from BH perturbation theory.
The merger-ringdown waveform, $h_{l m}^\textrm{merger-RD}$, is then stitched to
inspiral-plunge waveform, $h_{l m}^\textrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) =
h_{l m}^\textrm{insp-plunge}\, \Theta(t_\textrm{match}^{l m} - t)
+ h_{l m}^\textrm{merger-RD}\,\Theta(t-t_\textrm{match}^{l m})\,,
\end{equation}
%
where $\Theta(t)$ is the Heaviside step function \agcomm{Changed $\theta(t)$ to $\Theta(t)$,
because $\theta$ was already used to denote the parameter space.}.
%
The merger-ringdown waveform is expressed as an exponentially damped
sinusoid~\cite{Bohe:2016gbl,Cotesta:2018fcv,Mihaylov:2021bpf}
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})}\,,
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} = {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})
= \omega_{l m 0} - i / \tau_{l m 0}\,,
\end{align}
%
are the complex frequencies of the fundamental ($0$-th overtone) QNMs of the remnant BH.
%
\ag{Going back to the convention established above to use the index $k$ to now denote the $(l, m,0)$
mode,  the frequencies $\omega_k$ and damping times $\tau_k$ can be read off from the
real and imaginary parts of $\sigma_k$ respectively.}
%
The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined
in Ref.~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

In the $\SEOB$ model~\cite{Mihaylov:2021bpf}, the complex frequencies
$\sigma_k$ are computed by first determining the final mass and spin from
estimates of the initial masses and spins through NR-fitting-formulas~\cite{Taracchini:2013rva,Hofmann:2016yih}
and then converting them to the complex frequencies using BH perturbation
theory-inspired analytical fits~\cite{Berti:2005ys,Berti:2009kk}.
%
Hence,
%
\begin{subequations}
\begin{align}
\omega_k^{\text{GR}} &= \omega_k^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,
\\
\tau _k^{\text{GR}} &= \tau _k^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,
\end{align}
\end{subequations}
%
where $(\omega_k^{\text{GR}}, \tau_k^{\text{GR}} )$ refer to the
GR QNM predictions in the baseline \SEOB{} model.

\ag{In this paper, we replace these GR predictions with QNM
frequencies defined through the ParSpec expansion introduced in
Sec.~\ref{sec:review_parspec}; cf.~Eqs.~\eqref{eq:kerr_expansion}. Hence,}
%
\begin{subequations}
\begin{align}
\omega_k &= \omega_k(m_1, m_2, \chi_1, \chi_2,\ell, \{\delta \omega_k^{(n)}\}),\\
\tau_k   &= \tau _k(m_1, m_2, \chi_1, \chi_2, \ell, \{\delta \tau_k^{(n)}\}),
\end{align}
\end{subequations}
%
\ag{where the mass and spin of the remnant object $(M_f,\chi_f)$ are themselves functions of
$(m_1,\, m_2,\, \chi_1,\, \chi_2)$ (through~\cite{Taracchini:2013rva,Hofmann:2016yih}), and we fix $p$
to a certain theory-specific value. Additionally,  the frequencies depend on the ParSpec coefficients $\{\delta
\omega_k^{(n)},\delta \tau_k^{(n)}\}$,  and the coupling constant $\ell$.}

\sout{For example, if we restrict ourselves to corrections with $n=0$, or
$\chi_f^0$, to the $k=(2,2,0)$ mode only, then we would just have two
additional parameters, $\{\delta \omega_0^{(0)}, \delta \tau_0^{(0)}\}$.
%
In this paper, we will restrict ourselves to corrections to the $n=0$ and
$n=1$ terms, that is, only up to the leading-order in spin $\chi_f$ term.}

Using this parameterized waveform model, which we will call $\pSEOB$ in this
paper, we infer bounds on our non-GR parameter $\ell$ for specific cases of
modified gravity theories presented Sec.~\ref{sec:review_theories}.
%
We detail these results in Sec.~\ref{sec:results}.
%
% \hs{Alessandra, Abhirup, should we call this waveform differently to distinguish it from your previous papers?}
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

\subsection{General remarks on the interpretation of our results}
\label{sec:remarks}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
\ab{Some of these gravitational-wave events have sufficiently high
signal-to-noise ratio in the merger, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown, which is}
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we \sout{take a first step to} address this question by using an
inspiral-merger-ringdown waveform model in the effective-one-body formalism,
augmented with a parametrization of the ringdown based on an expansion \ab{in} the
final black hole's spin.
%
We give a prescription on how to include \ab{in} this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. \ab{In particular, we focus on gravity} theories that deviate
from general relativity through higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet, dynamical Chern-Simons, cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against \ab{the two loudest gravitational-wave ringdown signals, so far, 
  notably, GW150914 and GW200129. We find} that while the Einstein-dilaton-Gauss-Bonnet theory cannot be constrained with
these events, we can place upper bounds on the coupling constants of cubic-
($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order ($\ell_{\rm qEFT}
\leqslant 51.3$~km) effective-field-theories of general relativity, and
of dynamical Chern-Simons gravity ($\ell_{\rm CS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently
unconstrained by inspiral-only analyses which, however, can be constrained by
merger-ringdown studies \ab{with current gravitational-wave data}.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\ab{Unless otherwise specified, we will work in geometric units $G = 1 = c$.}

@@

\ab{Having reviewed the modified gravity theories that we will consider, we now present
the sequential building blocks for the waveform model we use to test these theories
against GW observations.
%
We start by reviewing the parametrized ringdown spin-expansion coefficients (ParSpec) framework in Sec.~\ref{sec:review_parspec}.
%
Next, in Sec.~\ref{sec:review_pSEOB}, we review our baseline
parametrized inspiral(--plunge)-merger-ringdown (IMR) waveform
model~\cite{Brito:2018rfr,Ghosh:2021mrv}, and explain how we extend
it to include the ParSpec.
%
Finally, in Sec.~\ref{sec:theory_specific_qnm}, we show how we can map
theory-specific QNM calculations in modified gravity theories onto the
free coefficients in the ParSpec framework.}
%
Ultimately, this provides us with an IMR waveform model, with the ringdown
portion of the model informed by QNM calculations in specific beyond-GR
theories.

@@

\ab{A general procedure to describe deviations to the QNM frequencies $\omega^{\rm GR}_{\ell m n}$ and
damping times $\tau^{\rm GR}_{\ell m n}$ of the GR BH is to write,}
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{\ell m n} &= \omega_{\ell m n}^{\rm GR} \, (1 + \delta \omega_{\ell m n}), \\
\tau_{\ell m n }   &= \tau_{\ell m  n}^{\rm GR}   \, (1 + \delta \tau_{\ell m n }),
\end{align}
\end{subequations}
%
\ab{where $\delta\omega_{\ell m n}$ and $\delta\tau_{\ell m  n}$ are the fractional deformation parameters,
$\ell$ and $m$ are the multipole indices, and $n$ the overtone number.}
%
This type of parametrization was adopted, for instance, in Refs.~\cite{Gossan:2011ha,Meidam:2014jpa,Brito:2018rfr,Carullo:2018sfu,Ghosh:2021mrv}.

The current LIGO-Virgo-KAGRA tests of BH ringdown (see Sec.VII.A
of Ref.~\cite{LIGOScientific:2020tif} or Sec.VIII.A
of Ref.~\cite{LIGOScientific:2021sio}) take a restrictive \ab{theory-independent} 
approach towards the inference of $\delta\omega_{\ell m n}$ and
$\delta\tau_{\ell m n}$. These deviations are either assumed to occur
identically across all observed sources or belong to a generic
underlying Gaussian population. A more realistic scenario, however, is
that these parameters depend on the source BH's mass and spin.
%
Ideally, one would like to explicitly reinstate this dependence, by
%
(i) introducing deformation parameters which can be determined, once and for
all, from a specific gravity theory (GR included), and
%
(ii) making it simpler to combine constraints coming from multiple (independently observed) sources.

\ab{The ParSpec framework was introduced in Ref.~\cite{Maselli:2019mjd} and can be used to our purpose. It is} an observable-based
bivariate expansion of Eq.~\eqref{eq:general_deviation}, given by
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{\ell m n} &= \frac{1}{M_{\rm f}} \sum_{j = 0}^{N_{\rm max}} \, \chi_{\rm f}^{j} \omega^{(j)}_{\ell m n} \, \left( 1 + \gamma\, \delta \omega^{(j)}_{\ell m n} \right), \\
\tau_{\ell m n}   &= M_{\rm f}     \sum_{j = 0}^{N_{\rm max}} \, \chi_{\rm f}^{j} \tau^{(j)}_{\ell m n}   \, \left( 1 + \gamma\, \delta \tau^{(j)}_{\ell m n} \right),
\end{align}
\end{subequations}
%
\ab{where $M_{\rm f}$ and $\chi_{\rm f}$ are the detector-frame final mass and spin, respectively;}
the quantities $\omega_{\ell m n}^{(j)}$ and $\tau_{\ell m n}^{(j)}$ are dimensionless coefficients of the
expansion in spin for BH QNMs in GR, while $\delta\omega_{\ell m n}^{(j)}$ and $\delta\tau_{\ell m n}^{(j)}$ are 
source-independent dimensionless coefficients that characterize the corrections to the GR QNM at
each spin-order,
%
and $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence for a given modified gravity theory is contained in the
dimensionless parameter $\gamma$, which reads
%
\begin{equation}
\gamma
%
= \left(\frac{\ell_{\rm th}}{M^{\rm s}_{\rm f}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M_{\rm f}}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on \ab{the length-scale parameter $\ell_{\rm th}$ of the specific gravity theory 
(non-GR modifications become important at distances $\lesssim \ell_{\rm th}$), and
the exponent $p$ is related to how the non-GR modifications are
added to the Einstein-Hilbert action.}
%
In Eq.~\eqref{eq:def_gamma} we made $\gamma$ dimensionless by using the other
length-scale associated with the remnant black hole (i.e., its source-frame mass
$M^{\rm s}_{\rm f}$\footnote{Throughout this paper, we maintain the convention of
using the upperscript ``s" to denote source-frame quantities and plain baseline
symbols for detector-frame measurements.}), which we can also write in terms of the detector-frame mass $M_{\rm f}$ 
through the redshift $z$~\cite{Krolak:1987ofj}. Also, dividing by the factor $G/c^2$ allows us to
express $\ell$ in physically intuitive metric units.

In principle, a modification to GR would also affect $M_{\rm f}$ and $\chi_{\rm f}$ and the
expansion should be written in terms of the non-GR mass and spin, say $\bar{M}_{\rm f}$ and
$\bar{\chi}_{\rm f}$. If we assume that the non-GR corrections are included perturbatively
(as it is the case with all the theories described in Sec.~\ref{sec:review_theories}),
the modifications to the BH mass and spin can be absorbed into the deviations
parameters $\delta\omega^{(j)}_{\ell m n}$ and $\delta\tau^{(j)}_{\ell m n}$.
%
This means we can identify the $M_{\rm f}$ and $\chi_{\rm f}$ with their \emph{corresponding {\rm GR} values}
\footnote{For a more detailed discussion, see Appendix A in Ref.~\cite{Maselli:2019mjd}.}.
%
We will see in Sec.~\ref{sec:pe}  that this assumption is indeed
satisfied in our parameter estimation studies (e.g., see Fig.~\ref{fig:corner_plot}).

Finally, we remark that in the GR limit ($\gamma \to 0$)
the series~\eqref{eq:kerr_expansion} truncated at $N_{\rm max} = 4$, reproduces with 1\% accuracy
the GR QNMs for BH's spins $\chi \lesssim 0.7$.
%
The values of the fitting coefficients $\omega_{\ell m n}^{(j)}$ and $\tau_{\ell m n}^{(j)}$
can be found in Ref.~\cite{Maselli:2019mjd}. 
%
In Ref.~\cite{Carullo:2021dui}, the fitting coefficients were calculated up to $N_{\rm max} = 9$,
which extend the validity of the spin-expansion up to $\chi \lesssim 0.99$.
%
As we will discuss in Sec.~\ref{sec:pe}, the expansion to $N_{\rm max} = 4$ is
sufficient for our purposes. \ab{For convenience we list the coefficients in the case of 
GR in Table~\ref{tab:ref_theories_qnms}.}

@@

We now describe the waveform model used in our paper to infer properties of a
BBH ringdown. 
%
As in Refs.~\cite{Brito:2018rfr,Ghosh:2021mrv}, we use an IMR BBH waveform
model where the \ab{(complex) frequencies} \sout{parameters} describing the remnant object 
are left additionally free and estimated directly from the data.

The GW signal from a \ab{quasicircular} BBH can be \ab{described} \sout{determined} in GR by a unique set of
parameters $\bm{\theta}$, that includes the masses and spins of the two BHs,
$(m_1,\, m_2,\, \mathbf{S}_1,\, \mathbf{S}_2)$, the sky location determined by the
luminosity distance $D_L$\abcomm{I cchanged to capital d because later we call the 
data ``d''.}, right ascension $\alpha$ and declination $\delta$,
and the orientation of the binary given by the inclination $\iota$ and polarization $\psi$
angles.
%
The set is completed by the choice of a reference time $t_0$ and phase
$\phi_0$. If we further assume that the spins of the individual BHs are
restricted to be \sout{parallel} \ab{aligned or anti-aligned (for short, aligned)} to the unit vector 
perpendicular to the orbital plane ($\mathbf{\hat{L}}$), we reduce the six
components of the spins to just two, $\chi_i \equiv \mathbf{S}_i\cdot \mathbf{\hat{L}} /m_i^2$ with $i = 1,2$, 
and our entire parameter set from 15 to 11.
%
Let us also define some additional parameters and set some conventions that
will be useful in our analysis later, namely, the total mass
%
$M=m_1+m_2$,
%
the chirp mass
%
$\mathcal {M}=(m_{1}m_{2})^{3/5}/(m_{1}+m_{2})^{1/5}$,
%
the asymmetric mass ratio $q=m_1/m_2$, with the convention $m_1 \geqslant m_2$ (and thus $q \geqslant 1$),
and the symmetric mass ratio of the binary, $\nu = m_1m_2/(m_1+m_2)^2$. \ab{Note that for BHs $-1 \leq \chi_i \leq 1$.}

For the polarizations of the GW signal (in the observer's frame) we have
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical
harmonics.
%
% --------------------------------
% HS: maybe not necessary anymore:
% \footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.}
% --------------------------------
%
As our baseline model, that is, the GR model upon which non-GR modifications
are added, we use the computationally efficient (time-domain) multipolar
waveform model for quasicircular spin-aligned BBHs described in
Ref.~\cite{Mihaylov:2021bpf}, which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$,
and $(5,5)$. \ab{Such a model was built by applying the post-adiabatic approximation~\cite{Nagar:2018gnk} 
to the multipolar spin-aligned effective-one-body (EOB) waveform model of Refs.~\cite{Bohe:2016gbl,Cotesta:2018fcv} 
(henceforth we refer to our baseline model as \SEOB{}~\footnote{This waveform
model is available in \texttt{LALSuite}~\cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA}
waveform approximant.}).}

%
An accurate description of the merger is incorporated through calibration with
NR simulations, as described in Refs.~\cite{Bohe:2016gbl,Cotesta:2018fcv}, along with
information for the merger and ringdown phases, from BH perturbation theory.
The merger-ringdown waveform, $h_{l m}^\textrm{merger-RD}$, is then stitched to
inspiral-plunge waveform, $h_{l m}^\textrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) =
h_{l m}^\textrm{insp-plunge}\, \Theta(t_\textrm{match}^{l m} - t)
+ h_{l m}^\textrm{merger-RD}\,\Theta(t-t_\textrm{match}^{l m})\,,
\end{equation}
%
where $\Theta(t)$ is the Heaviside step function. 
%
The merger-ringdown waveform is expressed as an exponentially damped
sinusoid~\cite{Bohe:2016gbl,Cotesta:2018fcv} 
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})}\,,
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} \equiv {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})
= \omega_{l m 0} - \frac{i}{\tau_{l m 0}}\,,
\end{align}
%
are the complex frequencies of the fundamental ($0$-th overtone) QNMs of the remnant BH.
%
%\ag{Going back to the convention established above to use the index $k$ to now denote the $(l, m,0)$
%mode,  the frequencies $\omega_k$ and damping times $\tau_k$ can be read off from the
%real and imaginary parts of $\sigma_k$ respectively.}
%
The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined
in Ref.~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

In the $\SEOB$ model~\cite{Bohe:2016gbl,Cotesta:2018fcv}, the complex frequencies
$\sigma_{\ell m 0}$ are computed by first determining the final mass and spin from
estimates of the initial masses and spins through NR-fitting-formulas~\cite{Taracchini:2013rva,Hofmann:2016yih}, 
and then converting them to the complex frequencies using BH perturbation
theory-inspired analytical fits~\cite{Berti:2005ys,Berti:2009kk}.
%
Hence,
%
\begin{subequations}
\begin{align}
\omega_{\ell m 0}^{\rm GR} &\equiv \omega^{\rm GR}_{\ell m 0}(m_1, m_2, \chi_1, \chi_2)\,,
\\
\tau _{\ell m 0}^{\rm GR} & \equiv \tau^{\rm GR}_{\ell m 0}(m_1, m_2, \chi_1, \chi_2)\,,
\end{align}
\end{subequations}
%
where $(\omega^{\rm GR}_{\ell m 0}, \tau^{\rm GR}_{\ell m 0})$ refer to the GR QNM predictions in the baseline \SEOB{} model. 
In this paper, we replace these GR predictions with QNM
frequencies defined through the ParSpec framework introduced in
Sec.~\ref{sec:review_parspec} (see Eqs.~\eqref{eq:kerr_expansion}). Hence,
%
\begin{subequations}
\begin{align}
\omega_{\ell m 0} & \equiv  \omega_{\ell m 0}(m_1, m_2, \chi_1, \chi_2,\ell_{\rm th}, \{\delta \omega_{\ell m 0}^{(j)}\}),\\
\tau_{\ell m 0}   & \equiv \tau_{\ell m 0} (m_1, m_2, \chi_1, \chi_2, \ell_{\rm th}, \{\delta \tau_{\ell m 0}^{(j)}\}),
\end{align}
\end{subequations}
%
where the mass and spin of the remnant object $(M_f,\chi_f)$ are themselves functions of
$(m_1,\, m_2,\, \chi_1,\, \chi_2)$~\cite{Taracchini:2013rva,Hofmann:2016yih}, and we fix $p$
to a certain theory-specific value. Additionally,  the frequencies depend on the ParSpec coefficients $\{\delta
\omega_{\ell m 0}^{(j)},\delta \tau_{\ell m 0}^{(j)}\}$,  and the \ab{length-scale} $\ell_{\rm th}$.

Using this parameterized waveform model, which we denote $\pSEOB$, 
we infer bounds on our non-GR parameter $\ell_{\rm th}$ for the specific cases of
modified gravity theories presented Sec.~\ref{sec:review_theories}.
%
We detail our results in Sec.~\ref{sec:results}.
%
% \hs{Alessandra, Abhirup, should we call this waveform differently to distinguish it from your previous papers?}
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

\subsection{Interpretation of our results}
\label{sec:remarks}
